---
title: "GAMtest"
output: html_document
---

```{r packages and data, cache=T}
require(sf)
require(raster)
require(sp)
require(tidyverse)
require(lubridate)
require(mgcv)
require(rnaturalearth)
require(lwgeom)

#wd <- "D:/Box Sync/Arctic/Data"
#setwd(wd)
# flickr data
#load("Flickr_Artic_60N_googlelabels_escodes_amap_plusPAs.Rdata")
load("D:/Box Sync/Arctic/Data/Flickr/processed/Flickr_Artic_60N_googlelabels_escodes_amap_plusPAs.Rdata")

# protected areas
PA <- st_read("Arctic_Protected_areas_2017/CAFF_Protected_Areas_20_01_2017_nplaea.shp")
#PA <- st_read("D:/Box Sync/Arctic/Data/Ecological/Arctic_Protected_areas_2017/CAFF_Protected_Areas_20_01_2017_nplaea.shp")
# because PA has some invalid geometries (i.e. st_is_valid()) we need to buffer PA to fix these
PAbuf <- st_buffer(PA,0)

# bounding box for entire Arctic
# bb <- st_sfc(st_multipoint(matrix(c(-180:180,rep(60,361)),ncol=2)),
#              crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ") %>%
#   st_sf() %>% 
#   st_transform(st_crs(flickramap)) %>% 
#   st_convex_hull()

#bb <- st_read("flickr_AMAP60N_dissolve.shp")
bb <- st_read("D:/Box Sync/Arctic/Data/Boundaries/Arctic_circle/AMAP/flickr_AMAP60N_dissolve.shp")

ggplot(bb)+
  geom_sf()

# country

# countries <- rbind(raster::getData('GADM', country='CAN', level=0),
#                    raster::getData('GADM', country='USA', level=0),
#                    raster::getData('GADM', country='ISL', level=0),
#                    raster::getData('GADM', country='NOR', level=0),
#                    raster::getData('GADM', country='FRO', level=0),
#                    raster::getData('GADM', country='FIN', level=0),
#                    raster::getData('GADM', country='GRL', level=0),
#                    raster::getData('GADM', country='SJM', level=0),
#                    raster::getData('GADM', country='SWE', level=0),
#                    raster::getData('GADM', country='RUS', level=0)) %>% 
#   st_as_sf() %>% 
#   st_transform(st_crs(flickramap)) %>% 
#   st_simplify(preserveTopology = TRUE,10000) #TODO:when doing for real, comment out this line!

# for natural earth, SJM seems to be lumped in woth NOR
countries <- ne_states(returnclass = "sf")  %>% 
  filter(adm0_a3 %in% c("CAN","USA","ISL","NOR","FRO","FIN","GRL","SWE","RUS")) %>% 
  st_transform(st_crs(flickramap)) %>% 
  lwgeom::st_make_valid() %>%
  st_intersection(bb) %>% 
  mutate(adm0_a3 = replace(adm0_a3, name=="Svalbard", "SJM")) %>% 
  group_by(adm0_a3) %>% summarise()
 
ggplot(countries)+
  geom_sf(aes(fill=adm0_a3))

# flickr
AllFlickr <- read.csv("D:/Box Sync/Arctic/Data/Flickr/global_flickr/Flickr_global_nphotostaken_byhr_2000to2018.csv") %>% 
#AllFlickr <- read.csv("Flickr_global_nphotostaken_byhr_2000to2018.csv") %>% 
  mutate(year=year(datetime)) %>% 
  filter(year>=min(flickramap$year)) %>% 
  group_by(year) %>% 
  summarize("Total Flickr Photos" = sum(total))

# naturalearth layers, change ne_load() with ne_download() if the data is not downloaded yet
roads <- ne_load(type="roads",
                     scale="large",
                     returnclass = "sf",
                     destdir = getwd())  %>% 
  st_transform(st_crs(flickramap)) %>% 
  st_intersection(bb)

ggplot(roads)+
  geom_sf()

airports <- ne_load(type="airports",
                     scale="large",
                     returnclass = "sf",
                     destdir = getwd())  %>% 
  st_transform(st_crs(flickramap)) %>% 
  st_intersection(bb)

ggplot(airports)+
  geom_sf()

ports <- ne_load(type="ports",
                     scale="large",
                     returnclass = "sf",
                     destdir = getwd())  %>% 
  st_transform(st_crs(flickramap)) %>% 
  st_intersection(bb)

ggplot(ports)+
  geom_sf()

populated_places <- ne_load(type="populated_places",
                     scale="large",
                     returnclass = "sf",
                     destdir = getwd())  %>% 
  st_transform(st_crs(flickramap)) %>% 
  st_intersection(bb)

ggplot(populated_places)+
  geom_sf()

urban_areas <- ne_load(type="urban_areas",
                     scale="large",
                     returnclass = "sf",
                     destdir = getwd())  %>% 
  st_transform(st_crs(flickramap)) %>% 
  st_intersection(bb)

ggplot(urban_areas)+
  geom_sf()

```

We have to create a hexagonal "raster" grid (not an actual raster)

```{r hex grid, cache=T}

makegridfun <- function(currdiameter){
    # switch to sp for the spsample function
    ext <- as(st_buffer(bb, currdiameter), "Spatial")
    projection(ext) <- sf::st_crs(flickramap)$proj4string
    
    # generate hexagon grid for footprints
    currgrid <- spsample(ext,
                     type = "hexagonal",
                     cellsize = currdiameter,
                     offset = c(0.5, 0.5)) %>% 
      HexPoints2SpatialPolygons(dx = currdiameter) %>% 
      st_as_sf() %>% 
      st_intersection(bb) %>% 
      rowwise() %>% 
      mutate(Latitude = st_coordinates(st_centroid(geometry))[2],
             Longitude = st_coordinates(st_centroid(geometry))[1]) %>% 
      st_as_sf(crs=st_crs(bb))
  return(currgrid)
}

grid_models <- makegridfun(100000) #we use a bigger resolution for the models
grid_footprint <- makegridfun(50000) #we use a bigger resolution for the models

```

Then, for the models, let's see which countries overlap which grid cell and if there is overlap with protected areas. If there is >20% overlap with PA then we consider that cell as a PA, we justify this because of the halo effect of PAs, people stay just outside PAs where there are accomodations.

```{r custom functions, cache=T}

# create a function to measure proportion of elements of sf1 are covered by sf2
prop_overlap <- function(sf1,sf2){
  area <- as.numeric(st_area(sf1))
  dist <- map(st_geometry(sf1),
              function(x) st_sfc(x,crs=st_crs(sf2)) %>% 
                st_intersection(sf2) %>% 
                st_area() %>% 
                as.numeric()) %>% 
    lapply(function(x) ifelse(is.null(x), NA, x)) %>% 
    lapply(function(x) ifelse(is.na(x), 0, x)) %>%
    unlist()
  return(dist/area)
}

# create a function to measure which elements of sf1 are most covered by elements of sf2
most_overlap <- function(sf1,sf2,column,missing){
  most <- suppressWarnings(map(st_geometry(sf1),
                               function(x) st_sfc(x,crs=st_crs(sf2)) %>% 
                                 st_sf() %>% 
                                 st_intersection(.,sf2) %>% 
                                 mutate(area=st_area(.)) %>%
                                 filter(area==max(area)) %>% 
                                 data.frame() %>% 
                                 select(column))) %>% 
    lapply(function(x) ifelse(nrow(x)==0, missing, x)) %>% 
    unlist()
  return(most)
}

# create a function to measure the length of sf2 contained in elements of sf1
length_overlap <- function(sf1,sf2){
  lengths <- map(st_geometry(sf1),
                 function(x) st_sfc(x,crs=st_crs(sf2)) %>% 
                   st_intersection(.,sf2) %>% 
                   st_combine() %>% 
                   st_sf() %>% 
                   mutate(length=as.numeric(sum(st_length(.)))) %>%
                   data.frame() %>% 
                   select(length)) %>% 
    lapply(function(x) ifelse(nrow(x)==0, 0, x)) %>% 
    unlist()
}
```

```{r overlaps,message=FALSE, dependson=c("packages and data", "custom functions", "hex grid"), cache=T}
# protected areas
grid_models$propPA <- prop_overlap(grid_models,PAbuf)
grid_models$PA <- grid_models$propPA>0.2
```

```{r overlaps1, message=FALSE, dependson=c("overlaps"), cache=T}
# country
grid_models$country <- most_overlap(grid_models,countries,"adm0_a3","Ocean")
```

```{r overlaps2, message=FALSE, dependson=c("overlaps1"), cache=T}
# roads
grid_models$roadlength <- length_overlap(grid_models,roads)
grid_models$dist2road <- st_distance(grid_models,st_combine(roads))
```

```{r overlaps3, message=FALSE, dependson=c("overlaps2"), cache=T}
# airports
grid_models$airports <- lengths(st_intersects(grid_models,airports))
grid_models$dist2airports <- st_distance(grid_models,st_combine(airports))
```

```{r overlaps4, message=FALSE, dependson=c("overlaps3"), cache=T}
# ports
grid_models$ports <- lengths(st_intersects(grid_models,ports))
grid_models$dist2ports <- st_distance(grid_models,st_combine(ports))
```

```{r overlaps5, message=FALSE, dependson=c("overlaps4"), cache=T}
# populated places
grid_models$populated_places <- lengths(st_intersects(grid_models,populated_places))
grid_models$dist2populated_places <- st_distance(grid_models,st_combine(populated_places))
```

```{r overlaps6, message=FALSE, dependson=c("overlaps5"), cache=T}
# urban area
grid_models$urban_areas <- prop_overlap(grid_models,urban_areas)
grid_models$dist2urban_areas <- st_distance(grid_models,st_combine(urban_areas))

head(grid_models)
save(grid_models, file = paste0("ArcticAMAP_griddedaccessibilitydata_",cell_diameter,"_m.Rdata"))
```

We also need to calculate the number of Flickr photos per year and correct it based on total Flickr usage over the years

```{r by year, cache=T, dependson="overlaps6"}

#add yearseason to flickramap
flickramap$season <- "summer"
flickramap$season[flickramap$month %in% c("01", "02", "03", "04", "11", "12")] <- "winter"
flickramap <- flickramap %>% mutate(yearseason = paste(year, season, sep="_"))
# 
# #  grid by yearseason
# gridYear <- sapply(split(flickramap, flickramap$yearseason),
#                    function(x) lengths(st_intersects(grid_footprint,x))) %>%
#   cbind(grid,.) %>%
#   gather(key=yearseason,value=photoCount,starts_with("X")) %>%
#   mutate(year=as.numeric(substr(strsplit(yearseason, "_")[[1]][1],2,5)),
#          season=strsplit(yearseason, "_")[[1]][2])

#photo-unit-day function
PUD_grid <- function(flickrrecords, time, grid){
  df <- flickrrecords %>% split(time) %>% 
    map( function(x){
           x$col.id <- as.integer(row.names(x))
           grid$row.id <- as.integer(row.names(grid))
           PUDdf <- as.data.frame(st_intersects(grid,x)) %>% 
             left_join(x,by="col.id") %>% 
             select(row.id,owner_date) %>% 
             unique() %>% 
             group_by(row.id) %>%
             summarize(PUD=n()) %>% 
             right_join(grid,by="row.id") %>% 
             mutate(PUD=if_else(is.na(PUD), 0, as.numeric(PUD))) %>%
             as.tibble() %>% select(row.id, PUD, Latitude, Longitude)
            return(PUDdf)
           })
  df <- do.call(rbind, df) %>% as.tibble
  return(df)
}

#The problem: This works, but it drops the time variable in the output 
#the output consists of a long row.id, PUD, Latitude, Longitude


     # gather(key=yearseason, value=PUD) %>% 
      #mutate(year=as.numeric(substr(strsplit(yearseason, "_")[[1]][1],2,5)),
     #    season=strsplit(yearseason, "_")[[1]][2])



#how many pud in each grid cell in each year-season, for the models
gridYearPUD_models <- PUD_grid(flickramap, flickramap$yearseason, grid_models) 

#how many photos in each grid cell in each year-season, for the footprint analysis
gridYearPUD_footprint <- PUD_grid(flickramap, flickramap$yearseason, grid_footprint) 



byYear <- gridYearPUD %>%
  data.frame() %>% 
  group_by(yearseason) %>% 
  summarise("Artic Photos"=sum(photoCount)) %>% 
  left_join(AllFlickr,by="yearseason") %>% 
  mutate("Percent Traffic"=.$"Artic Photos"/.$"Total Flickr Photos",
         correctionFact=mean(.$"Total Flickr Photos")/.$"Total Flickr Photos") %>% 
  gather(key="variable",value="value",-yearseason)

correct <- byYear %>% 
  filter(variable=="correctionFact") %>% 
  select(-variable)
byYear <- byYear %>% 
  filter(variable!="correctionFact")

save(gridYear,file = paste0("gridYear_",cell_diameter,"_m.Rdata"))
save(gridYearPUD,file = paste0("gridYearPUD_",cell_diameter,"_m.Rdata"))

```

Now that we have all our basic data ready to go, let's vizualize!

First, lets see if there is a change in the number of users over time?
```{r users, cache=T}
plotData <- byYear
#plot showing all, arctic, and % traffic
unique(plotData$variable)
#reorder
plotData$variable <- ordered(plotData$variable,levels=c("Total Flickr Photos","Artic Photos","Percent Traffic"))
ggplot(plotData) + 
  geom_line(aes(x=yearseason,y=value)) +
  facet_grid(vars(variable),scales="free_y")
```

Both the total number of photos on Flickr and in the Arctic increase over time, but the Arctic represents an increasing share of Flickr's yearly photo traffic.

Are these increasing numbers of tourists always visiting the same places, or are they exploring new grounds?

```{r footprint, cache=T}
# spatial footprint expanded?
plotData <- filter(gridYear, PUD>0) %>% 
  left_join(correct,by="yearseason") %>% 
  mutate(correctedPUD=photoCount*value)

ggplot(plotData)+
  geom_histogram(aes(x=correctedPUD))+
  scale_y_log10() +
  facet_wrap(vars(yearseason, season), ncol=2) 
```

```{r footprint2, cache=T}
# or
plotData <- gridYearPUD %>% 
  group_by(yearseason) %>% 
  summarise(footprint=mean(photoCount>0)) %>% 
  mutate(type="Overall Flickr Footprint")

# this samples n records from each year in flickramap
# n was the number of records in the first year
# we want an n equal to the smallest year*season combo, so min(table(flickramap$yearseason))
equaln_sample_flickr <- flickramap %>% 
  group_by(yearseason) %>% 
  # sample_n(as.numeric(table(flickramap$year)[1])) #before
  sample_n(min(table(flickramap$yearseason)))


# this recalculates PUD by yearseason by grid cell for equaln_sample_flickr
gridYear_equaln <- PUD_grid(equaln_sample_flickr,equaln_sample_flickr$yearseason,grid) %>% 
  cbind(grid,.) %>% 
  gather(key=yearseason,value=PUD,starts_with("X")) %>% 
  mutate(year=as.numeric(substr(strsplit(yearseason, "_")[[1]][1],2,5)),
         season=strsplit(yearseason, "_")[[1]][2])

plotData_equaln <- gridYear_equaln %>% 
  group_by(yearseason) %>% 
  summarise(footprint=mean(PUD>0) )%>% 
  mutate(type="With Equal N")


trafficn_sample_flickr <- flickramap %>% 
  # mutate(year=as.numeric(year)) %>% 
  nest(-yearseason)  %>% 
  left_join(correct,by="yearseason") %>% 
  mutate(n=round(map_dbl(data, nrow)/max(value)*value)) %>%
  mutate(Sample = map2(data, n, sample_n)) %>%
  unnest(Sample) %>% 
  st_as_sf(crs = st_crs(flickramap))

table(trafficn_sample_flickr$yearseason)
table(flickramap$yearseason)

gridYear_trafficn <- sapply(split(trafficn_sample_flickr, trafficn_sample_flickr$yearseason),
                          function(x) lengths(st_intersects(grid,x))) %>% 
  cbind(grid,.) %>% 
  gather(key=yearseason,value=PUD,starts_with("X")) %>% 
  mutate(year=as.numeric(substr(strsplit(yearseason, "_")[[1]][1],2,5)),
         season=strsplit(yearseason, "_")[[1]][2])

plotData_trafficn <- gridYear_trafficn %>% 
  group_by(yearseason) %>% 
  summarise(footprint=mean(PUD>0)) %>% 
  mutate(type="With Increased Arctic Traffic")

plotDataAll <- rbind(plotData,plotData_equaln,plotData_trafficn)

plotDataAll$type <- ordered(plotDataAll$type,levels = unique(plotDataAll$type)[c(1,3,2)])

ggplot(plotDataAll)+
  geom_area(aes(x=yearseason,y=footprint,fill=type),position = "identity")+
  scale_fill_manual(values=c("#1f78b4","#b2df8a","#a6cee3"))+
  facet_wrap(vars(season))
```

```{r footprint3, cache=T}
# or
plotData <- gridYear %>% 
  group_by(year,country) %>% 
  summarise(footprint=mean(photoCount>0))

ggplot(plotData)+
  geom_line(aes(x=year,y=footprint,color=country))
```

Let's see if we can see what is driving these changes. First lets model whether people visit protected areas more than expected

```{r GAM, cache=T, dependson="by year"}

gridYearmod <- gridYear %>% 
  filter(country!="Russia") %>% 
  mutate(photoCountlog10=log10(photoCount+1))

#set norway as the reference level
gridYearmod <- within(gridYearmod, country <- relevel(country, ref = "Norway"))
#dropped year, dropped populatedplaces, dist2popplaces, urbanareas
g <- gam(photoCountlog10 ~ s(Latitude)+
           s(Longitude)+
           country+
           propPA+
           season+
           PA+
           roadlength+
           dist2road+
           airports+
           dist2airports+
           ports+
           dist2ports+
           dist2urban_areas,
         data = gridYearmod, method = "REML")
gam.check(g)
summary(g)
```

Now lets see if people use different types of access in different seasons.

```{r GAM_seasonal, cache=T, dependson="by year"}

gridYearsummermod <- gridYear %>% 
  filter(country!="Russia") %>% 
  filter(season=="summer") %>% 
  mutate(photoCountlog10=log10(photoCount+1))

gridYearwintermod <- gridYear %>% 
  filter(country!="Russia") %>% 
  filter(season=="winter") %>% 
  mutate(photoCountlog10=log10(photoCount+1))

#Summer
gridYearsummermod <- within(gridYearsummermod, country <- relevel(country, ref = "Norway"))
gs <- gam(photoCountlog10 ~ s(Latitude)+
           s(Longitude)+
           year+
           country+
           propPA+
           PA+
           roadlength+
           dist2road+
           airports+
           dist2airports+
           ports+
           dist2ports+
           populated_places+
           dist2populated_places+
           urban_areas+
           dist2urban_areas,
         data = gridYearsummermod, method = "REML")
gam.check(gs)
summary(gs)
AIC(gs)

#Winter
gridYearwintermod <- within(gridYearwintermod, country <- relevel(country, ref = "Norway"))
gw <- gam(photoCountlog10 ~ s(Latitude)+
           s(Longitude)+
           year+
           country+
           propPA+
           PA+
           roadlength+
           dist2road+
           airports+
           dist2airports+
           ports+
           dist2ports+
           populated_places+
           dist2populated_places+
           urban_areas+
           dist2urban_areas,
         data = gridYearwintermod, method = "REML")
gam.check(gw)
summary(gw)
```
